---
title: "Summary of camera trap data, SAGW, 2022"
date: "2023-05-26"
output: 
  word_document:
    reference_docx: "../../../word-styles-01.docx"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Effort

Information about when cameras were deployed and retrieved 

Information about sampling occasions (number, dates for each, proportion of time that cameras were operational?)

Include a figure with camera deployment information, with sampling occasions denoted?

# Detections

Table with detection summaries

- row for each species detected (common and scientific names)
- number of photos, detections (max of one per day or sampling period per location)
- number of locations
- detection rate (denoting for which species we ran occupancy model)

# Modeling approach

Occupancy models

- types of data used
- parameters estimated

spOccupancy package (that uses a Bayesian framework)

Model selection

- what covariates were considered (i.e., candidate model set)
- how a model for inference was selected

# Species 1 (create species sections in a loop since the number of species will change across parks and years?)

## Model used for inference

Brief description

Table with parameter estimates (along with generic caption)


Example:
```{r cars}
summary(cars)
```

## Estimated occurrence probabilities

Map

## Estimated covariate effects on occurrence and/or detection probabilities

Figures with captions. Example: 

```{r pressure, echo=FALSE}
par(mar = c(4, 4, 1, 1))
plot(pressure)
```

