---
params: 
  date: "2023-06-20"
  PARK: "SAGW"
  YEAR: 2022    
title: "Summary of camera trap data, `r params$PARK`, `r params$YEAR`"
date: "`r params$date`"
output:
  officedown::rdocx_document:
    reference_docx: "word-styles-02.docx"
---


```{r setup, include = FALSE}

library(lubridate)
library(tidyverse)
library(terra)
library(spOccupancy)
library(tidyterra)
library(officer)
library(officedown)
library(flextable)

knitr::opts_knit$set(root.dir = "../" )
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dev = "ragg_png",
  fig.align = "left",
  fig.width = 5, 
  fig.height = 3.5, 
  fig.dpi = 150, 
  fig.cap.style = "Caption",
  fig.cap.pre = "Figure ",
  fig.cap.sep = ". ",
  fig.cap.fp_text = fp_text_lite(bold = TRUE),
  tab.align = "left",
  tab.cap.style = "Caption",
  tab.cap.pre = "Table ",
  tab.cap.sep = ". ",
  tab.cap.fp_text = fp_text_lite(bold = TRUE))

set_flextable_defaults(
  font.family = "Arial", 
  font.size = 9, 
  big.mark = "", 
  table.layout = "autofit")

tab_num = run_autonum(seq_id = "table", bkm = NULL)
fig_num = run_autonum(seq_id = "figure", bkm = NULL)

```

```{r data-formatting}

source("src/functions.R")
source("src/photo-data/format-mammal-data.R")

PARK <- params$PARK
YEAR <- params$YEAR

```

# Effort
```{r effort}

# Park long name
PARKL <- ifelse(PARK == "SAGW", 
                "the Tucson Mountain District of Saguaro National Park",
                ifelse(PARK == "CHIR", "Chiricahua National Monument",
                       "Organ Pipe Cactus National Monument"))

# Subset events data and add a numeric camera ID
events <- events %>%
  filter(Park == PARK, d_yr == YEAR) %>%
  mutate(locnum = as.numeric(as.factor(StdLocName))) %>%
  as.data.frame()
deploy_start <- format(min(events$d_date), "%d %B")
deploy_end <- format(max(events$d_date), "%d %B")
retr_start <- format(min(events$r_date), "%d %B")
retr_end <- format(max(events$r_date), "%d %B")

# Load sampling occasion data
occasions <- read.csv("data/occasions/occasions-all-parks.csv")
occasions <- occasions %>%
  filter(Park == PARK) %>%
  filter(yr %in% YEAR) %>%
  mutate(start_yday = yday(start),
         end_yday = yday(end),
         mid_yday = round((start_yday + end_yday)/2)) %>%
  arrange(yr, occasion) 
season_start <- format(as.Date(occasions$start[1]), "%d %B")
season_end <- format(as.Date(occasions$end[nrow(occasions)]), "%d %B")

# Function to create chart with deployments
ggcust_segs <- function(dat, park, year, xlims, occ){
  ggplot() +
    geom_segment(filter(dat, Park == park & d_yr == year),
                 mapping = aes(x = d_date, xend = r_date, y = locnum, yend = locnum),
                 linewidth = 0.3, color = "dodgerblue3") + 
    geom_segment(filter(dat, Park == park & d_yr == year),
                 mapping = aes(x = d_date, xend = d_date, y = locnum+0.4, yend = locnum-0.4),
                 linewidth = 0.3, color = "dodgerblue3") + 
    geom_segment(filter(dat, Park == park & d_yr == year),
                 mapping = aes(x = r_date, xend = r_date, y = locnum+0.4, yend = locnum-0.4),
                 linewidth = 0.3, color = "dodgerblue3") +
    geom_vline(xintercept = as.Date(occ$start), color = "gray50") + 
    geom_vline(xintercept = as.Date(occ$end[nrow(occ)]), color = "gray50") +
    labs(x = "", y = "Camera number") + 
    scale_x_date(limits = xlims, date_labels = "%b") +
    theme(text=element_text(size = 10),
          axis.title.x=element_blank(),
          plot.margin = margin(0.1, 0.2, 0.2, 0.2, unit = "cm"))
}

xmin <- paste0(YEAR, "-01-01")
if (PARK == "SAGW") {
  xmax <- paste0(YEAR, "-03-15")
} else {
  xmax <- paste0(YEAR, "-12-31")
}

segs <- ggcust_segs(events, PARK, YEAR, xlims = as.Date(c(xmin, xmax)),
                    occ = occasions)
```

A total of `r nrow(events)` cameras were deployed in SAGW between 
`r deploy_start` and `r deploy_end` and were retrieved between 
`r retr_start` and `r retr_end`, `r YEAR`. We delineated a total of 
`r nrow(occasions)` sampling occasions that were each `r occasions$duration[1]` 
days long.

```{r deployment-table}

occ_short <- occasions %>%
  select(c(occasion, start, end)) %>%
  rename(Occasion = occasion,
         Start = start,
         End = end)
occ_ftable <- flextable(occ_short)
occ_ftable <- flextable::align(occ_ftable, align = "left", part = "all")

occ_ftable <- set_caption(occ_ftable,
                          caption = paste0("Start and end dates for each sampling occasion in ",
                                 PARKL, ", ", YEAR, "."),
                          autonum = tab_num)
occ_ftable <- bold(occ_ftable, bold = TRUE, part = "header")

```

```{r tab-deploy, ft.align = "left", tab.id = "table"}

occ_ftable

```

<br>

```{r deployment-plot, fig.id = "figure", fig.cap = segs_cap}

segs_cap <- paste0("Camera deployments and sampling occasions in ", PARKL, ", ",
  YEAR, ". Each horizontal blue line represents the period over which a ",
  "camera was operational, from the date the camera was deployed through the ", 
  "date the camera was retrieved. Gray vertical lines denote the beginning ",
  "and end of the ", nrow(occasions), " consecutive sampling occasions.")

segs

```


# Detections
```{r detections}

# Clean up species table
species <- species %>%
  select(-n)
species$Species[species$Species_code == "SKUNK"] <- "Mephitidae"
species$Species[species$Species_code == "UNCA"] <- "Canidae"

# Filter detection data
dat <- dat %>%
  filter(Park == PARK & yr == YEAR) %>%
  filter(obsdate >= occasions$start[1] & 
           obsdate <= occasions$end[nrow(occasions)])
nphotos <- dat %>%
  group_by(Species_code) %>%
  summarize(n_photos = length(datetime),
            n_locs = length(unique(LocationName))) %>%
  arrange(desc(n_photos)) %>%
  left_join(species, by = "Species_code") %>%
  relocate(Common_name:Species, .before = n_photos) %>%
  data.frame()

# Look at detection data for various species
# Where detection is max of one per sampling occasion at each location
detects <- read.csv("output/species-detections-byparkyr.csv", header = TRUE)
detects <- detects %>%
  dplyr::filter(Park == PARK & yr == YEAR) %>%
  select(c(spp, ndetects)) %>%
  rename(Species_code = spp,
         n_detects = ndetects)

nphotos <- nphotos %>%
  left_join(detects, by = "Species_code") %>%
  relocate(n_detects, .after = n_photos) %>%
  select(-Species_code)

```

We detected a total of `r nrow(nphotos)` mammal species on the `r nrow(events)`
cameras during the `r nrow(occasions)` sampling occasions. 

```{r detections-table}

nphotos_cap <- paste0("Mammal species detected in ",
  PARKL, " between ", season_start, " and ", season_end, ", ", YEAR, 
  ". No. photos is the total number of photographs obtained (multiple photos ",
  "may occur at the same camera location in the same day). No. detections is ",
  "the number of detections that could be used in an occupancy modeling ",
  "framework (maximum of one photographic detection per camera location per ",
  "sampling period). No. locations is the unique number of camera locations ",
  "where the species was photographed.")

nphotos_ft <- flextable(nphotos) %>%
  align(j = 1:2, align = "left") %>%
  align(j = 3:5, align = "center") %>%
  italic(j = 2, italic = TRUE, part = "body") %>%
  bold(bold = TRUE, part = "header") %>%
  set_header_labels(Common_name = "Common name",
                    Species = "Scientific name",
                    n_photos = "No. photos",
                    n_detects = "No. detections",
                    n_locs = "No. locations") %>%
  set_caption(caption = nphotos_cap, autonum = tab_num)

```

```{r tab-detects, ft.align = "left", tab.id = "table"}

nphotos_ft

```


# Modeling approach

## Occupancy models

```{r data-prep}

# Get list of species that we ran occupancy models for (and list of files with 
# model output)
spp_rds <- list.files(path = "output/single-season-models/", 
                      pattern = paste0(PARK, "-", YEAR),
                      full.names = FALSE)
spp_table <- data.frame(Species_code = str_sub(spp_rds, 11, 14))
spp_table$common <- species$Common_name[match(spp_table$Species_code,
                                              species$Species_code)]
nspp <- nrow(spp_table)
spp_code <- spp_table$Species_code

# Extract MCMC parameters from one of these model objects
model_list <- readRDS(paste0("output/single-season-models/", spp_rds[1]))
n_samples <- model_list$model$n.samples
n_thin <- model_list$model$n.thin
n_burn <- model_list$model$n.burn
n_chains <- model_list$model$n.chains
n_post <- model_list$model$n.post

# Load multi-layer raster with spatial data for park
park_raster <- readRDS(paste0("data/covariates/spatial-cov-", PARK, ".rds"))
# We have two distance-to-boundary layers, one that applies to the entire park
# boundary and one that applies to boundaries that are adjacent to unprotected
# lands (boundaryUP). For now, we'll remove the original boundary layer.
park_raster <- subset(park_raster, "boundary", negate = TRUE)
names(park_raster)[names(park_raster) == "boundaryUP"] <- "boundary"

# Extract dataframe with covariate values at each camera location
spatial_covs <- model_list$data$occ.covs

# Load general information about covariates
covariates <- read.csv("data/covariates/covariates.csv")

```

For each species with a sufficient number of detections (here, `r nspp` 
species), we used single-season occupancy models to estimate the probability of 
occurrence and the probability of detection given occurrence (MacKenzie et al. 
2002). To estimate these parameters, we generated encounter histories for each 
camera location, where a "1" denotes that the species was detected at least once 
during a sampling occasion and a "0" indicates that the species was not 
detected. For example, an encounter history of "10011" would indicate that a 
species was photographed at least once during the first, fourth, and fifth 
sampling occasions and was not photographed during the second and third sampling 
occasions. 

We used a Bayesian framework to estimate model parameters, as this made it 
easier to estimate derived parameters (e.g., proportion of area occupied across 
the entire park) with associated uncertainties and to incorporate random effects 
that can account for uncertainties beyond that explained by covariates in the 
model. We fit models in R using the spOccupancy package (Doser et al. 2022). For 
each model, we ran `r n_chains` Markov chains initiated at random values for 
`r n_samples` iterations. We discarded the first `r n_burn` iterations and 
retained 1 of every `r n_thin` iterations thereafter, using the remaining 
`r n_chains * n_post` samples (across all the chains) to summarize the posterior 
distribution.

```{r spp-loop, results = "asis"}

modelfile_base <- paste0(getwd(), "/output/single-season-models/")
knitchild_base <- paste0(getwd(), "/src/")

child <- lapply(spp_code, function(x) {
  SPECIES <- x
  modelfile <- paste0(modelfile_base, PARK, "-", YEAR, "-",  SPECIES, ".rds")
  model_list <- readRDS(modelfile)
  knitr::knit_child(paste0(knitchild_base, "single-season-report-spp.Rmd"),
                    envir = environment(),
                    quiet = TRUE)
})
cat(unlist(child), sep = "\n")

```

