
```{r}

spp_name <- spp_table$common[spp_table$Species_code == SPECIES]
cap_exceptions <- c("American", "Arizona", "North", "Sonoran")
spp_name <- paste(sapply(unlist(str_split(spp_name, " ")),
                         function(x) {ifelse(x %in% cap_exceptions,
                                             x,
                                             tolower(x))}),
                  collapse = " ")
spp_plural <- ifelse(str_detect(string = spp_name, 
                                pattern = "sheep|deer|pronghorn|elk"),
                     spp_name, ifelse(str_detect(string = spp_name,
                                                 pattern = "fox"),
                                      paste0(spp_name, "es"), 
                                      paste0(spp_name, "s")))

best <- model_list$model
psi_model <- model_list$psi_model
p_model <- model_list$p_model
data_list <- model_list$data

# Extract names of covariates (with and without "_z" subscripts) from best model
psi_covs_z <- create_cov_list(psi_model)
if (length(psi_covs_z) == 1 & any(psi_covs_z == "1")) {
  psi_covs_z <- character(0)
  psi_covs <- character(0)
} else {
  psi_covs <- psi_covs_z %>% str_remove_all(pattern = "_z")
}
p_covs_z <- create_cov_list(p_model)
if (length(p_covs_z) == 1 & any(p_covs_z == "1")) {
  p_covs_z <- character(0)
  p_covs <- character(0)
} else {
  p_covs <- p_covs_z %>% str_remove_all(pattern = "_z")
}

# Create table with summary stats that can be saved to file
occ_estimates <- parameter_estimates(model = best,
                                     parameter = "occ",
                                     lower_ci = 0.025,
                                     upper_ci = 0.975)
det_estimates <- parameter_estimates(model = best,
                                     parameter = "det",
                                     lower_ci = 0.025,
                                     upper_ci = 0.975)
occ_estimates <- occ_estimates %>%
  rename(Covariate = Parameter) %>%
  mutate(Parameter = "Occurrence", .before = "Covariate")
det_estimates <- det_estimates %>%
  rename(Covariate = Parameter) %>%
  mutate(Parameter = "Detection", .before = "Covariate")
estimates <- rbind(occ_estimates, det_estimates)

# Calculate the number of covariates in each part of the model:
n_covs_occ <- length(psi_covs)
n_covs_occt <- ifelse(n_covs_occ > 1, paste0(n_covs_occ, " covariates"), 
                      ifelse(n_covs_occ == 1, "1 covariate", "no covariates"))
n_covs_det <- length(p_covs)
n_covs_dett <- ifelse(n_covs_det > 1, paste0(n_covs_det, " covariates"), 
                      ifelse(n_covs_det == 1, "1 covariate", "no covariates"))

```

# Output for `r SPECIES`

## Model used for inference

The highest-ranking model for `r spp_plural` included `r n_covs_occt` in the 
occurrence part of the model and `r n_covs_dett` in the detection part of the
model.

```{r}

estimates <- estimates %>%
  rename(LowerCI = "Lower95%",
         UpperCI = "Upper95%") %>%
  mutate(across(Mean:UpperCI, function(x) round(x, 2))) %>%
  mutate("95% CI" = paste0(LowerCI, ", ", UpperCI)) %>%
  select(-c(Median, LowerCI, UpperCI, exclude0)) %>%
  relocate("95% CI", .after = SD)

estimates_cap <- paste0("Parameter estimates (on the logit scale) from a ",
  "model for ", spp_plural, " in ", PARKL, ", ",  YEAR, ". SD = Standard ",
  "deviation; 95% CI = 95% credible interval. Rhat values between 1 and 1.05 ",
  "indicate that the model has converged. ESS = effective sample size; values ",
  "> 400 are usually sufficient. f values indicate the proportion of posterior ",
  "samples that are < 0 if the mean is < 0 or the proportion of samples that ",
  "are > 0 if the mean is > 0. All continuous covariates were standardized by ",
  "their respective means and standard deviations prior to analysis.")

estimates_ft <- flextable(estimates) %>%
  align(j = 1:2, align = "left", part = "all") %>%
  align(j = 3:8, align = "center", part = "all") %>%
  bold(bold = TRUE, part = "header") %>%
  colformat_double(j = c(3:4, 6, 8), digits = 2) %>%
  set_caption(caption = estimates_cap, autonum = tab_num) %>%
  add_footer_lines("")

```

```{r, ft.align = "left", tab.id = "table"}

estimates_ft

```

## Estimated detection probabilities

```{r}

alpha_samples <- best$alpha.samples

# Calculate overall detection probability (for experienced deployment personnel
# if that was included in the model)
if ("deploy_exp" %in% p_covs) {
  alpha_deploy <- alpha_samples[ ,c("(Intercept)", "deploy_exp")]
  X0 <- as.vector(c(1, 2))
  probs_logit <- alpha_deploy %*% X0
} else {
  probs_logit <- alpha_samples[ ,"(Intercept)"]
}
det_probs <- exp(probs_logit) / (1 + exp(probs_logit))
mean_det <- mean(det_probs)
lcl_det <- quantile(det_probs, 0.025)
ucl_det <- quantile(det_probs, 0.975)

# Right now, we can't produce figures for day or effort effect because 
# we don't day/effort values on the original scale in model_list$data
# We'd need to add this to the model_list somewhere.

```

Overall detection probability for `r spp_plural` `r if("deploy_exp" %in% p_covs) 
{"with experienced deployment personnel "}`was `r sprintf("%.2f", mean_det)` 
(95% credible interval [CI] = `r sprintf("%.2f", lcl_det)`, 
`r sprintf("%.2f", ucl_det)`). 

## Estimated occurrence probabilities

```{r}

beta_samples <- best$beta.samples

# Calculate overall occurrence probability (or for each vegetation class if 
# that's in the model)
if ("vegclass2" %in% psi_covs) {
  vegclass_table <- vegclass_estimates(model_list$model,
                                       parameter = "occ",
                                       lower_ci = 0.025,
                                       upper_ci = 0.975) %>%
    mutate(across(ci_lower:ci_upper, function(x) round(x, 2))) %>%
    mutate("95% CI" = paste0(ci_lower, ", ", ci_upper)) %>%
    select(-c(ci_lower, ci_upper)) %>%
    relocate("95% CI", .after = sd_prob)
  vegocc_cap <- paste0("Mean occurrence estimates for ", spp_plural, 
    " for each vegetation class in ", PARKL, ", ",  YEAR, ".")
  vegocc_ft <- flextable(vegclass_table) %>%
    align(j = 1, align = "left") %>%
    align(j = 2:4, align = "center", part = "all") %>%
    bold(bold = TRUE, part = "header") %>%
    colformat_double(j = 2:3, digits = 2) %>%
    set_caption(caption = vegocc_cap, autonum = tab_num) %>%
    set_header_labels(vegclass = "Veg class",
                      mean_prob = "Mean",
                      sd_prob = "SD") %>%
    add_footer_lines("")
  mean_occ <- NULL
  lcl_occ <- NULL
  ucl_occ <- NULL
} else {
  probs_logit <- beta_samples[ ,"(Intercept)"]
  occ_probs <- exp(probs_logit) / (1 + exp(probs_logit))
  mean_occ <- mean(occ_probs)
  lcl_occ <- quantile(occ_probs, 0.025)
  ucl_occ <- quantile(occ_probs, 0.975)
  vegocc_ft <- NULL
}

```

```{r, include = ("vegclass2" %in% psi_covs), ft.align = "left", tab.id = "table"}

vegocc_ft

```

```{r, include = (!"vegclass2" %in% psi_covs), results = "asis"}

res <- knitr::knit_child(text = paste0("Overall occurrence probability for ",
                                       spp_plural, " was ", 
                                       sprintf("%.2f", mean_occ), " (", 
                                       sprintf("%.2f", lcl_occ), ", ", 
                                       sprintf("%.2f", ucl_occ), ")."),
                         envir = globalenv())
cat(unlist(res))

```

```{r code = readLines("src/single-season-models/spOccupancy-predictions.R"), include = length(psi_covs) > 0, eval = TRUE}

```

```{r, include = length(psi_covs) > 0, fig.id = "figure", fig.cap = occmap_cap, eval = TRUE}

occmap_cap <- paste0("Predicted probability of ", spp_name, " occurrence in ", 
                     PARKL, " in ", YEAR, ". ")
occmap_base <- ggplot() +
  geom_spatraster(data = preds_mn, mapping = aes(fill = mean)) +
  scale_fill_viridis_c(na.value = "transparent") +
  labs(fill = "", x = NULL, y = NULL, title = NULL) +
  theme_bw()
occmap_base + theme(axis.title = element_blank(), 
                    title = element_blank())

```

## Covariate effects

```{r}

# Identify continuous covariates in occurrence part of the model
  psi_continuous <- psi_covs_z[!psi_covs_z %in% c("1", "vegclass2", "vegclass3")]
  psi_cont_unique <- unique(psi_continuous)
  psi_n_cont <- length(psi_cont_unique)

# If there are any continuous covariates, create a figure for each:
  if (psi_n_cont > 0) {
    # Loop through each covariate
    for (cov in psi_cont_unique) {
      # Create name of plot:
      plotname <- paste0("marginal_psi_", str_remove(cov, "_z"))
      # Create plot
      assign(plotname, 
             marginal_plot_occ(covariate = cov, 
                               model = best, 
                               data_list = data_list,
                               covariate_table = covariates,
                               central_meas = mean))
    } 
  }

  marg_figs <- str_subset(ls(), "marginal_psi_")

```

```{r, include = length(marg_figs) > 0, fig.id = "figure", fig.cap = margfig_cap}

for (marg_fig in marg_figs) {
  cov_name <- str_remove(marg_fig, "marginal_psi_")
  cov_name <- ifelse(cov == "burn_severity_2011",
                     "burn",
                     str_remove(cov, "_z"))
  cov_name <- tolower(covariates$axis_label[covariates$short_name == cov_name])
  margfig_cap <- paste0("Effect of ", cov_name, 
                        " on occurrence probability of ", spp_plural, " in ", 
                        PARKL, ", ", YEAR, ".")
  print(get(marg_fig))
}

```



