---
params: 
  date: "2023-06-20"
  PARK: "SAGW"
  YEAR: 2023    
output:
  word_document:
    reference_docx: SR_Report_Style.docx
    df_print: paged
---


```{r setup, include = FALSE}

library(lubridate)
library(tidyverse)
library(terra)
library(spOccupancy)
library(tidyterra)
library(officer)
library(officedown)
library(flextable)

knitr::opts_knit$set(root.dir = "../" )
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  dev = "ragg_png",
  fig.align = "left",
  fig.width = 5, 
  fig.height = 3.5, 
  fig.dpi = 150, 
  fig.cap.style = "Figure caption",
  fig.cap.pre = "Figure ",
  fig.cap.sep = ". ",
  fig.cap.fp_text = fp_text_lite(bold = TRUE),
  tab.align = "left",
  tab.cap.style = "Table Caption",
  tab.cap.pre = "Table ",
  tab.cap.sep = ". ",
  tab.cap.fp_text = fp_text_lite(bold = TRUE))

set_flextable_defaults(
  font.family = "Arial", 
  font.size = 9, 
  big.mark = "", 
  table.layout = "autofit")

tab_num = run_autonum(seq_id = "table", bkm = NULL)
fig_num = run_autonum(seq_id = "figure", bkm = NULL)

Fig_alt_text <- ":::{custom-style='sr Alternate text'}\n\nFigure X. Alt Text\n\n:::\n\n"

Fig_credit <- ":::{custom-style='sr Image credit'}\n\nNPS (update as needed)\n\n:::\n\n"

```

```{r data-formatting}

PARK <- params$PARK
YEAR <- params$YEAR
# Park long name
PARKL <- ifelse(PARK == "SAGW", 
                "the Tucson Mountain District of Saguaro National Park",
                ifelse(PARK == "CHIR", "Chiricahua National Monument",
                       "Organ Pipe Cactus National Monument"))

title_var <- paste("Wildlife Monitoring at", PARKL,"–",YEAR)

date2 <- as.character(Sys.Date())


source("src/photo-data/format-mammal-data.R")
source("src/functions.R")

# Load sampling occasion data
occasions <- read.csv(paste0("data/occasions/occasions-", PARK, ".csv"))
```

:::{custom-style="sr Title"}
`r title_var`
:::

:::{custom-style="sr Image line"}
Add your cover image here.
:::
:::{custom-style="sr Cover image caption"}
Cover Image Caption
:::
:::{custom-style="sr Alternate text"}
  Cover IMAGE Alt Text
:::
:::{custom-style="sr Cover image credit"}
© COVER IMAGE CREDIT
::: 

  
FirstName MiddleInitial. LastName (ORCID if available)^1^, FirstName MiddleInitial. LastName(ORCID if available)^2^, Erin Zylstra (ORCID 0000-0002-2536-0403)^3\*,4^
  
^1^ National Park Service, Chihuahuan Desert Network,  
3655 Research Dr., Genesis Building D,  
Las Cruces, NM 88003
  
^2^ National Park Service, Sonoran Desert Network,  
12661 E. Broadway Blvd.,  
Tucson, AZ 85748

^3^ Tucson Bird Alliance
300 E University Blvd. #120
Tucson, AZ 85705

^4^ USA National Phenology Network
1311 E 4th St
Tucson, AZ 85721

^5^ Organization Name, Office/Unit Name 2 (optional) (NPS Office/Unit, University Dept., etc.), Street Address, City Name, ST ZipCode (zip is optional)

^\*^ formerly

\newpage
# Abstract
You must write an abstract in paragraph format (400 words or less). You will also need to choose which figures you want to include, number them, write alt text for them, and reinsert the png into the Word document (to get higher resolution). You should also ensure that tables and figures are formatted properly.

This report was generated on `r date2`. Use the text and outputs below to populate the text highlighted in yellow and the sections that tell you tocopy from the Rmd output in the Annual Report template on Teams. In general, the yellow in the overview and key points can be populated once you've finished the results section.

We may or may not want to include all outputs for all species. The idea is that someone looks at the outputs and selects examples to include in the report. That may mean focusing on a few species or a few themes. Or it may mean including all species but not showing all tables/figures for each species. You will also need to produce publication quality graphics for your figures of interest. 

__Keywords:__ mammals, monitoring, wildlife, wildlife cameras, wildlife monitoring, occupancy, `r PARKL`

\newpage
# Executive Summary
Use the Annual Report template on Teams

## Key points
Use the Annual Report template on Teams

\newpage
# Background
Use the Annual Report template on Teams

## Occupancy modeling
Use the Annual Report template on Teams

# Methods
Use the Annual Report template on Teams and integrate the information below.

## Field methods
Use the Annual Report template on Teams

## Photo identification
Use the Annual Report template on Teams

## Single-species single-season occupancy models

```{r data-prep}

# Get list of species that we ran occupancy models for (and list of files with 
# model output)
spp_rds <- list.files(path = "output/single-season-models/", 
                      pattern = paste0(PARK, "-", YEAR),
                      full.names = FALSE)
species_rds <- str_remove(spp_rds, ".rds")
spp_table <- data.frame(Species_code = str_sub(species_rds, 
                                               11, 
                                               nchar(species_rds)))
spp_table$common <- species$Common_name[match(spp_table$Species_code,
                                              species$Species_code)]
nspp <- nrow(spp_table)
spp_code <- spp_table$Species_code

# Extract MCMC parameters from one of these model objects
model_list <- readRDS(paste0("output/single-season-models/", spp_rds[1]))
n_samples <- model_list$model$n.samples
n_thin <- model_list$model$n.thin
n_burn <- model_list$model$n.burn
n_chains <- model_list$model$n.chains
n_post <- model_list$model$n.post

# Load multi-layer raster with spatial data for park
park_raster <- readRDS(paste0("data/covariates/spatial-cov-", PARK, ".rds"))
# We have two distance-to-boundary layers, one that applies to the entire park
# boundary and one that applies to boundaries that are adjacent to unprotected
# lands (boundaryUP). For now, we'll remove the original boundary layer.
park_raster <- subset(park_raster, "boundary", negate = TRUE)
names(park_raster)[names(park_raster) == "boundaryUP"] <- "boundary"

# Extract dataframe with covariate values at each camera location
spatial_covs <- model_list$data$occ.covs

# Load general information about covariates
covariates <- read.csv("data/covariates/covariates.csv")

```

For each species with a sufficient number of detections (here, `r nspp` 
species), we used single-season occupancy models to estimate the probability of 
occurrence and the probability of detection given occurrence (MacKenzie et al. 
2002). First, we defined the season as those days between the first deployment date and 
the last retrieval date when ≥ 60% of the cameras were operational. We then delineated
consecutive 7-day sampling occasions within the season, with the first sampling occasion
beginning on the first day ≥ 60% of the cameras were operational. For each species and 
camera location, we generated encounter histories, where a "1" 
denoted that the species was detected at least once during a sampling occasion and a "0"
indicated that the species was not detected. For example, an encounter history of "10011"
would indicate that a species was photographed at least once during the first, fourth, and fifth 
sampling occasions and was not photographed during the second and third sampling 
occasions. 

We used a Bayesian framework to estimate model parameters, as this made it 
easier to estimate derived parameters (e.g., proportion of area occupied across 
the entire park) with associated uncertainties. We fit models in R using the 
spOccupancy package (Doser et al. 2022). For each model, we ran `r n_chains` 
Markov chains initiated at random values for `r n_samples` iterations. We 
discarded the first `r n_burn` iterations (the burn in) and retained 1 of every `r n_thin` 
iterations thereafter, using the remaining `r n_chains * n_post` samples (across 
all the chains) to summarize the posterior distribution. We verified that the effective sample sizes for model parameters were > 400.

## Covariates and model selection

```{r covariates-by-park}

topo <- "topographic (aspect, elevation, slope)"
anthro <- "human-caused (distance to nearest road, park boundary, trail, point-of-interest)"
veg_related <- "vegetation-related (vegetation classes and distance to nearest wash)"
burn_related <- "burn-related (burn severity during 2011 fire)"
lta_related <- "landtype association-related (broad landtype classes)"

cov_list <- ifelse(PARK == "SAGW",
                   paste0(topo, ", ", veg_related, ", and ", anthro),
                   ifelse(PARK == "ORPI",
                          paste0(topo,", ", lta_related, " and ", anthro),
                          paste0(topo, ", ", burn_related, ", and ", anthro)))

```

For each species, we first evaluated whether day of the year (using linear and quadratic terms), 
effort (proportion of days during each sampling occasion that the camera was 
operational), and the experience of deployment personnel influenced detection probability. 
Personnel were categorized as novice (<1 year experience), experienced (1-2 years experience) 
or expert (2+ years experience). We identified a number of `r cov_list` spatial covariates for `r PARKL` that 
could explain variation in occurrence probabilities. We then assembled a set of candidate 
models where each model included the best detection covariate(s) and one element from 
each occupancy covariate category, excluding any combination of covariates that were 
highly correlated (pairwise correlation coefficient, |r| > 0.7). We used widely applicable 
information criterion (WAIC) to identify the best supported candidate model, and then removed any covariates 
with little to no explanatory power (*f* < 0.9; where *f* represents the proportion of posterior samples that were < 0 if the posterior 
mean was < 0 or the proportion of posterior samples > 0 if the posterior mean was > 0 ) ). 
When necessary to estimate spatial variation in occupancy
or detection, we included camera location as a random effect. 
We used the resulting model for all inferences.

## Data and code availability
Use the Annual Report template on Teams

# Results 

## Effort
```{r effort}

# Subset events data and add a numeric camera ID
events <- events %>%
  filter(Park == PARK, d_yr == YEAR) %>%
  mutate(locnum = as.numeric(as.factor(loc))) %>%
  as.data.frame()
deploy_start <- format(min(events$d_date), "%B %d")
deploy_end <- format(max(events$d_date), "%B %d")
retr_start <- format(min(events$r_date), "%B %d")
retr_end <- format(max(events$r_date), "%B %d")

# count number of each camera type
cam_type <- events %>% group_by(CameraType) %>% count() %>% rename(count=n) %>% arrange(count) %>% mutate(CamText = paste(count, CameraType))
cam_type_text <- ifelse(cam_type$count[1] == sum(cam_type$count), as.character(cam_type$CameraType[1]), paste(unlist(cam_type$CamText), collapse =", "))

# Occasions by year
occasions_yr <- occasions %>%
  group_by(Park, yr) %>%
  count()

# Format sampling occasion data
#occasions <- read.csv(paste0("data/occasions/occasions-", PARK, ".csv"))
occasions <- occasions %>%
  filter(yr %in% YEAR) %>%
  mutate(start_yday = yday(start),
         end_yday = yday(end),
         mid_yday = round((start_yday + end_yday)/2)) %>%
  arrange(yr, occasion) 
season_start <- format(as.Date(occasions$start[1]), "%B %d")
season_end <- format(as.Date(occasions$end[nrow(occasions)]), "%B %d")

# Function to create chart with deployments
ggcust_segs <- function(dat, park, year, xlims, occ){
  ggplot() +
    geom_segment(filter(dat, Park == park & d_yr == year),
                 mapping = aes(x = d_date, xend = r_date, y = locnum, yend = locnum),
                 linewidth = 0.3, color = "dodgerblue3") + 
    geom_segment(filter(dat, Park == park & d_yr == year),
                 mapping = aes(x = d_date, xend = d_date, y = locnum+0.4, yend = locnum-0.4),
                 linewidth = 0.3, color = "dodgerblue3") + 
    geom_segment(filter(dat, Park == park & d_yr == year & r_date < xlims[2]),
                 mapping = aes(x = r_date, xend = r_date, y = locnum+0.4, yend = locnum-0.4),
                 linewidth = 0.3, color = "dodgerblue3") +
    geom_vline(xintercept = as.Date(occ$start), color = "gray50") + 
    geom_vline(xintercept = as.Date(occ$end[nrow(occ)]), color = "gray50") +
    labs(x = "", y = "Camera number") + 
    scale_x_date(limits = xlims, date_labels = "%d-%b") +
    theme(text=element_text(size = 10),
          axis.title.x=element_blank(),
          plot.margin = margin(0.1, 0.2, 0.2, 0.2, unit = "cm"))
}

xmin <- as.Date(occasions$start[1]) - 15
xmax <- as.Date(occasions$end[nrow(occasions)]) + 15

events_plotting <- events %>%
  mutate(r_date = if_else(r_date > xmax, xmax, r_date))

segs <- ggcust_segs(events_plotting, PARK, YEAR, xlims = as.Date(c(xmin, xmax)),
                    occ = occasions)
```

A total of `r nrow(events)` cameras (`r cam_type_text`) were deployed in `r PARKL` between 
`r deploy_start` and `r deploy_end` and were retrieved between 
`r retr_start` and `r retr_end`, `r YEAR`. We delineated a total of 
`r nrow(occasions)` sampling occasions that were each `r occasions$duration[1]` 
days long. The first sampling occasion began on `r season_start` and the last sampling occasion ended on `r season_end`. Community scientists, park staff, and/or partner organizations typically assist with camera deployments and retrievals. In `r YEAR`,  __##__ volunteers,  __##__ interns, and  __##__ park staff helped deploy and retrieve cameras in `r PARKL`. We are grateful for their support.

```{r deployment-table}

occ_short <- occasions %>%
  select(c(occasion, start, end)) %>%
  rename(Occasion = occasion,
         Start = start,
         End = end)

occtable_cap <- paste0("Start and end dates for each sampling occasion in ",
                       PARKL, ", ", YEAR, ".")

occ_ftable <- flextable(occ_short) %>%
  align(align = "left", part = "all") %>%
  bold(bold = TRUE, part = "header") %>%
  set_caption(caption = occtable_cap, autonum = tab_num) %>%
  add_footer_lines("")

```

```{r tab-deploy, ft.align = "left", tab.id = "table"}

occ_ftable

```

```{r deployment-plot, fig.id = "figure", fig.cap = segs_cap}

segs_cap <- paste0("Camera deployments and sampling occasions in ", PARKL, ", ",
  YEAR, ". Each horizontal blue line represents the period over which a ",
  "camera was operational, from the date the camera was deployed through the ", 
  "date the camera was retrieved. Gray vertical lines denote the beginning ",
  "and end of the ", nrow(occasions), " consecutive sampling occasions.")

segs

```

## Detections
```{r detections}

# Clean up species table
species <- species %>%
  select(-c(TSN, Family, Nativeness)) 
species$Species[species$Species_code == "UNFO"] <- "Vulpes spp. or Urocyon spp."
species$Species[species$Species_code == "UNLE"] <- "Lepus spp."
species$Species[species$Species_code == "UNOD"] <- "Odocoileus spp."
species$Common_name[species$Species_code == "UNCA"] <- "unknown canid"

# Filter observation data
# total photos during deployment 
total_photos <- (dat %>%
  filter(Park == PARK & yr == YEAR) %>%
  group_by(Park, yr) %>%
  count())$n

# any new species observation
obs_new <- dat %>%
  mutate(visit = ifelse(yr == YEAR, "Current", ifelse(yr > YEAR, "Future", "Past"))) %>%
  filter(visit != "Future") %>%
  ungroup() %>%
  group_by(Park, Species_code, visit) %>%
  summarize(sumobs = n(), .groups="keep") %>%
  pivot_wider(., names_from="visit", values_from="sumobs", values_fill=0) %>%
  filter(Past==0 & Current > 0) %>%
  left_join(species, by = "Species_code") %>%
  mutate(name = paste0(Common_name," (_",Species,"_)", sep=""))
  

obs_new_run = ifelse(PARK == "ORPI", FALSE, ifelse(nrow(obs_new) > 0, TRUE, FALSE))

# within park and within given years' occasion start & end
dat <- dat %>%
  filter(Park == PARK & yr == YEAR)
nphotos <- dat %>%
  group_by(Species_code) %>%
  summarize(n_photos = length(datetime),
            n_locs = length(unique(loc))) %>%
  data.frame()
dat <- dat %>%
  filter(obsdate >= occasions$start[1] & 
           obsdate <= occasions$end[nrow(occasions)])

# Look at detection data for various species
# Where detection is max of one per sampling occasion at each location
detects <- read.csv(paste0("output/species-detections-byyr-", PARK, ".csv"))

# any new detections
detects_new <- detects %>%
  mutate(visit = ifelse(yr == YEAR, "Current", ifelse(yr > YEAR, "Future", "Past"))) %>%
  filter(visit != "Future") %>%
  ungroup() %>%
  group_by(Park, spp, visit) %>%
  summarize(sumdetects = sum(ndetects), .groups="keep") %>%
  pivot_wider(., names_from="visit", values_from="sumdetects", values_fill=0) %>%
  filter(Past==0 & Current > 0) %>%
  left_join(species, by = c("spp" = "Species_code")) %>%
  mutate(name = paste0(Common_name," (_",Species,"_)", sep=""))

#new_species_run = ifelse(PARK == "ORPI", FALSE, ifelse(nrow(obs_new) > 0 | nrow(detects_new) > 0, TRUE, FALSE))

new_species_obs_text = ifelse(nrow(obs_new) > 0, paste("In ",YEAR, ", our cameras captured ",paste(unlist(obs_new$name), collapse=", "), " for the first time during a SODN camera deployment.", sep=""), "")

new_species_det_text = ifelse(nrow(detects_new) > 0, paste("In ",YEAR, ", we detected ",paste(unlist(detects_new$name), collapse=", "), " for the first time during a sampling occasion.", sep=""), "")
                              
new_species_text <- ifelse(new_species_obs_text == "" & new_species_det_text == "", "All species listed in Table 1 were observed and detected during previous SODN monitoring.", paste(new_species_obs_text, new_species_det_text))


# for selected year
detects <- detects %>%
  dplyr::filter(yr == YEAR) %>%
  select(c(spp, ndetects)) %>%
  rename(Species_code = spp,
         n_detects = ndetects)

nphotos <- nphotos %>%
  left_join(detects, by = "Species_code") %>%
  arrange(desc(n_photos)) %>%
  left_join(species, by = "Species_code") %>%
  relocate(Common_name:Species, .before = n_photos) %>%
  relocate(n_detects, .after = n_photos) %>%
  select(-Species_code)

# summarize species & genera counts for observations 
species_type_obs <- nphotos %>%
  mutate(separate_wider_delim(., cols=Species, delim=" ", names=c("Genus", "SpEp"), too_few = "align_start", too_many="drop")) %>%
  mutate(Type = ifelse(is.na(SpEp),"Genera","Species")) %>%
  mutate(Type = ifelse(Common_name == "unknown animal", "Unknown", Type))

# count species and genera separately
species_count_obs <- nrow(filter(species_type_obs, Type == "Species"))
genera_count_obs <- nrow(filter(species_type_obs, Type == "Genera"))

species_text_obs <- ifelse(genera_count_obs > 0, paste("We observed a total of ", species_count_obs, " mammal species plus ", genera_count_obs, " mammalian families or genera", sep=""), paste("We observed a total of ", species_count_obs, " mammal species", sep=""))

# summarize species & genera counts for detections 
species_type_det <- nphotos %>%
  filter(n_detects > 0) %>%
  mutate(separate_wider_delim(., cols=Species, delim=" ", names=c("Genus", "SpEp"), too_few = "align_start", too_many="drop")) %>%
  mutate(Type = ifelse(is.na(SpEp),"Genera","Species")) %>%
  mutate(Type = ifelse(Common_name == "unknown animal", "Unknown", Type))

# count species and genera separately
species_count_det <- nrow(filter(species_type_det, Type == "Species"))
genera_count_det <- nrow(filter(species_type_det, Type == "Genera"))

species_text_det <- ifelse(species_count_obs == species_count_det & genera_count_obs == genera_count_det & genera_count_obs == 0, paste("with at least one detection of each of the ", species_count_obs, " species", sep=""), ifelse(species_count_obs == species_count_det & genera_count_obs == genera_count_det & genera_count_obs > 0, paste("with at least one detection of each of the ", species_count_obs, " species and ", genera_count_obs, " genera/families", sep=""), ifelse(species_count_obs > species_count_det & genera_count_obs == 0, paste("but only had detections for ", species_count_det, " of the ", species_count_obs, " species", sep=""), ifelse(species_count_obs > species_count_det & genera_count_obs == species_count_det, paste("but only had detections for ", species_count_det, " of the ", species_count_obs, " species but all ",genera_count_obs, " of the genera/families", sep=""), ifelse(species_count_obs == species_count_det & genera_count_obs > genera_count_det, paste("with at least one detection of each of the ", species_count_obs, " species but only had detections for ", genera_count_det, " of the ", genera_count_obs, " genera/families", sep=""),"hmmm")))))

```

`r species_text_obs` in `r total_photos` photos from the `r nrow(events)` cameras during the `r YEAR` camera deployment. During the `r nrow(occasions)` sampling occasions (weeks) in `r YEAR`, we recorded `r nrow(dat)` detections `r species_text_det` (Table 1). 
`r {new_species_text}`

**Take a look and see if there are any interesting or notable detections, even if there weren’t many of them. If there were new observations or detections, those will be mentioned above.**


```{r detections-table}

nphotos_cap <- paste0("Mammal species detected in ",
  PARKL, " in ", YEAR, 
  ". No. photos is the total number of photographs obtained (multiple photos ",
  "may occur at the same camera location in the same day) in ", YEAR, 
  ", regardless of date. No. detections is ",
  "the number of detections that could be used in an occupancy modeling ",
  "framework (maximum of one photographic detection per camera location per ",
  "sampling occasion). No. locations is the unique number of camera locations ",
  "where the species was photographed.") 

nphotos_ft <- flextable(nphotos %>% dplyr::select(-Protected)) %>%
  align(j = 1:2, align = "left") %>%
  align(j = 3:5, align = "center") %>%
  italic(j = 2, italic = TRUE, part = "body") %>%
  bold(bold = TRUE, part = "header") %>%
  set_header_labels(Common_name = "Common name",
                    Species = "Scientific name",
                    n_photos = "No. photos",
                    n_detects = "No. detections",
                    n_locs = "No. locations") %>%
  set_caption(caption = nphotos_cap, autonum = tab_num) %>%
  add_footer_lines("")

```

```{r tab-detects, ft.align = "left", tab.id = "table"}

nphotos_ft

```

## Single-species, single-season occupancy
```{r spp-loop, results = "asis"}

modelfile_base <- paste0(getwd(), "/output/single-season-models/")
knitchild_base <- paste0(getwd(), "/src/")

child <- lapply(spp_code, function(x) {
  SPECIES <- x
  modelfile <- paste0(modelfile_base, PARK, "-", YEAR, "-",  SPECIES, ".rds")
  model_list <- readRDS(modelfile)
  knitr::knit_child(paste0(knitchild_base, "single-season-report-spp.Rmd"),
                    envir = environment(),
                    quiet = TRUE)
})
cat(unlist(child), sep = "\n")

```


\newpage
# Literature Cited
Beery, S., D. Morris, and S. Yang. 2020. Efficient Pipeline for Camera Trap Image Review [Computer software]. http://github.com/agentmorris/MegaDetector  

Brashares, J. S. 2010. Filtering wildlife. Science 329:402.

Carroll, C. 2010. Role of climatic niche models in focal-species-based conservation planning: Assessing potential effects of climate change on Northern Spotted Owl in the Pacific Northwest, USA. Biological Conservation 143:1432–1437.

Chen, C., and others. 2022. Global camera trap synthesis highlights the importance of protected areas in maintaining mammal diversity. Conservation Letters 15(2).

Doser, J. W., A. O. Finley, M. Kéry, and E. F. Zipkin. 2022. spOccupancy: An R package for single-species, multi-species, and integrated spatial occupancy models. Methods in Ecology and Evolution 13:1670–1678.

Hubbard, J. A., E. C. Dillingham, and A. Buckisch. In prep. Terrestrial mammal monitoring protocol implementation plan for the Sonoran Desert Network.

`r if(PARK!="SAGW"){"Hubbard, J. A., C. L. McIntyre, S. E. Studd, T. Nauman, D. Angell, K. Beaupre, B. Vance, and M. K. Connor. 2012. Terrestrial vegetation and soils monitoring protocol and standard operating procedures: Sonoran Desert and Chihuahuan Desert networks, version 1.1. Natural Resource Report NPS/SODN/NRR—2012/509. National Park Service, Fort Collins, Colorado."}`

Ivan, J. S., and E. S. Newkirk. 2016. CPW Photo Warehouse: A custom database to facilitate archiving, identifying, summarizing and managing photo data collected from camera traps. Methods in Ecology and Evolution 7:499–504.
MacKenzie, D. I., and J. A. Royle. 2005. Designing occupancy studies: general advice and allocating survey effort. Journal of Applied Ecology 42:1105-1114.

MacKenzie, D. I., J. D. Nichols, G. B. Lachman, S. Droege, J. Andrew Royle, and C. A. Langtimm. 2002. Estimating site occupancy rates when detection probabilities are less than one. Ecology 83:2248–2255.

MacKenzie, D. I., J. D. Nichols, J. E. Hines, M. G. Knutson, and A. B. Franklin. 2003. Estimating site occupancy, colonization, and local extinction when a species is detected imperfectly. Ecology 84:2200–2207. 

McIntyre C.L., M. Christensen, A. Buckisch, E. Dillingham, T. Atella, and J. McGeverly J. 2025. Mammal data from PARK, Arizona: year-year - internal. National Park Service. Fort Collins CO 

Post, E., J. Brodie, M. Hebblewhite, A. D. Anders, J. A. K. Maier, and C. C. Wilmers. 2009. Global population dynamics and hot spots of response to climate change. Bioscience 59:489–497.

Steenweg, R., J. Whittington, and M. Hebblewhite. 2015. Canadian Rockies Remote Camera Multi-Species Occupancy Project: Examining trends in carnivore populations and their prey using remote cameras in Banff, Jasper, Kootenay, Yoho and Waterton Lakes National Parks. Final Report. University of Montana. 88p. 

`r if(PARK!="SAGW"){"Theobald, D. M., D. L. Stevens, Jr., D. White, N. S. Urquart, A. R. Olsen, and J. B. Norman. 2007. Using GIS to generate spatially balanced designs for natural resource applications. Environmental Management 40:134–146."}`

Zylstra, E. 2025. SWParks_Mammal_Occupancy repository, new-data-exports branch. Available online at https://github.com/ezylstra/SWParks_Mammal_Occupancy/tree/new-data-exports

\newpage
# Appendix A
